Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2020-04-30 docs/00-manuscripts/ieee/ieee-manuscript.tex
don't try to compare SVM performance to other studies

\subsubsection{Model differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).

FIXME: Add refs when SVM was clearly worse then other algs.
In other (environmental) studies in recent years, SVM showed most often less good performances and was not able to outperform competing algorithms such as RF.
The difference to the second best performing model of around three percentage points is also worth noting:
We are not aware of a study in which SVM ranked as the best model that clearly.
All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).

@@

\subsection{Performance vs.\ plot characteristics}
\label{subsec:perf-plot-char}

The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).

>>>>

\subsubsection{Model differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
Comparing these differences (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).

@@

\subsection{Performance vs.\ plot characteristics}
\label{subsec:perf-plot-char}
The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
